APPM report - Draft

### 2022-09-11 paper_version/main.tex
basic setup

\documentclass{report}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{nicematrix}
\usepackage{subcaption}
\usepackage{pdfpages}
\usepackage{algpseudocode}
\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{listings}
\lstset { %
    language=C++,
    backgroundcolor=\color{white}, % set backgroundcolor
    numbers=none,
    frame=shadowbox,
    basicstyle=\footnotesize,% basic font setting
}

\NiceMatrixOptions{cell-space-top-limit = 4pt}
\NiceMatrixOptions{cell-space-bottom-limit = 4pt}

\usepackage{bm}
\usepackage{appendix}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem*{remark}{Remark}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tcolorbox}
\tcbuselibrary{theorems}
\newtcbtheorem
  []% init options
  {tcb-def}% name
  {Definition}% title
  {%
    colback=green!5,
    colframe=green!35!black,
    fonttitle=\bfseries,
  }% options
  {def}% prefix
\newtcbtheorem
  []% init options
  {tcb-prop}% name
  {Proposition}% title
  {%
    colback=blue!5,
    colframe=blue!80!black,
    fonttitle=\bfseries,
  }% options
  {prop}% prefix

\usepackage{natbib}
\usepackage[figurename=Fig.,tablename=Tab.,labelfont={bf}]{caption}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, decorations.markings, arrows, intersections}
\bibliographystyle{apalike}
\setcitestyle{authoryear,open={(},close={)}} %Citation-related commands
\usepackage[linkbordercolor=green]{hyperref}
\usepackage{geometry}
\geometry{a4paper,scale=0.8}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\newcommand{\mathcolorbox}[2]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\note}[1]{{\footnotesize \textcolor{blue}{[\textbf{#1}]}}}

\title{\textbf{Asymptotic-preserving Plasma Model in 3D}}
\author{Tianwei Yu}
\date{\today}

\begin{document}

\begin{titlepage}
\noindent \includegraphics[]{ETHlogo.pdf}

\centering
\vspace*{7cm}

\textbf{\LARGE Asymptotic-Preserving Plasma Model in 3D}

\vspace{3cm}
\large

Master thesis

\vspace{0.25cm}

\textbf{Tianwei Yu}

\vspace{0.25cm}

\today

\vspace{7cm}

Supervisor: Prof. Dr. Ralf Hiptmair

\vspace{0.25cm}

Department of Mathematics, ETH Z\"{u}rich

\end{titlepage}

\chapter*{Abstract}
Plasma physics involves complex phenomena characterized by multiple scales, for example, the propagation of electromagnetic fields, the interaction between charged particles and electromagnetic fields, and the drift of fluids. To numerically simulate plasmas, discretization schemes that can automatically adapt to the scale of interest and remain stable regardless of the system scales are desirable. In this thesis, we discuss asymptotic-preserving (AP) schemes for the Euler-Maxwell system parametrized by the dimensionless Debye length $\lambda$ in a 3D setting for multi-species plasma modeling. The key capability of the AP scheme is the uniform performance, in the sense of stability and convergence, for a series of problems that are parametrized by arbitrary $\lambda$ and constitute a singular perturbation as $\lambda \rightarrow 0$. As the first part of the thesis, the AP property, still being a vague concept throughout literature, is clarified by two aspects: asymptotic-stability and asymptotic-consistency, which respectively refer to the unconditional stability on the asymptotic parameter and the convergence of the limit scheme to the limit model. Two propositions discussing some sufficient conditions for general AP schemes are presented, hopefully providing deeper insights into the AP property and some guidance for devising AP schemes. In the sequel, we reproduce the results in \cite{degond_2012}, based on which we extend the 1D two-fluid discrete Euler-Maxwell model to a more general model in a 3D setting that involves multiple species, friction terms, and possibly other source terms. The 3D discrete model for the Euler-Maxwell system is implemented based on the frameworks of Finite Volume Method (FVM) and Finite Integration Technique (FIT) for the Euler system and the Maxwell system respectively. Details about the procedure of time-stepping and the assembly of the linear system due to implicitness are provided. An essential aspect lies in the choice of right boundary conditions that are compatible with both non-limit and limit models. The discrete model is tested on a case that is adapted from \cite{fuchs_2021} due to the difficulty arising from the insulating solid domain. As verification, the scheme is checked on the subsystems and is compared with the corresponding 1D discrete model. The numerical results on the fully-coupled system with $\lambda$ ranging in $[0,1]$ manifest a continuous transition from the non-neutral model ($\lambda = \mathcal{O}(1)$) to the quasi-neutral model ($\lambda = 0$), which, as well as the convergence of the solutions over time, implies that the scheme is AP.

\chapter*{Acknowledgements}
First, I would like to extend my deepest gratitude to my supervisor Prof. Ralf Hiptmair. His proposal for this thesis combines fluid dynamics, a topic that I am familiar with, and electromagnetism, a completely new world for me. With reasonable amounts of theoretical and implementation work, I had much fun in the process of freely exploring the knowledge and accomplishing the task step by step. His constant and insightful input was always inspiring at the moment when I was lost. In a word, it was a pleasant, adventurous, and memorable journey.

My appreciation also goes to Dr. Roman Fuchs, who gave me a lot of guidance on familiarizing myself with the topic and the underlying C++ code. I would also like to acknowledge the help of my colleague Langwen Huang in the configuration of the running environment on the cluster, which potentially saved me considerable time.

Meanwhile, I would like to thank my girlfriend, friends, and flatmates for their company and support during the thesis work as well as my Master's life. I wish them all a bright future.

In the end, I would like to express my special gratitude to my parents for their financial support and emotional care. They are always my strongest back no matter what setbacks I encounter.

\tableofcontents
\chapter{Introduction}
\section{Plasma}
Often described as the fourth state of matter, plasma is an ionized gas phase substance that consists of positively-charged \emph{ions}, negatively-charged \emph{electrons} and neutrally-charged atoms. Although plasmas seem to be intangible in our daily life, most of the material in the universe is in the plasma state \citep{gurnett2005}. They are largely associated with various astrophysical phenomena like sunspots, solar flares, planetary radiation belts. Regarding terrestrial plasmas, they can be found or exploited in many situations, for instance, lightning, gas-discharge lamps, controlled thermonuclear fusion.

A plasma is classically defined as \emph{a quasi-neutral gas of charged particles showing collective behavior} \citep{chen_1994}. The \emph{collective} behavior arises due to the long-range nature of $1/r$ Coulumb potential \citep{gibbon2020}. Besides, the dynamics of plasma particles are largely determined by an internally-driven electromagnetic field, in the sense that the charges generate an electric field, and carry an electric current that induces a magnetic field. In contrast, a gas in a neutral state is largely characterized by inter-particle collision. The second key feature, \emph{quasi-neutrality}, refers to the tendency to keep the overall charges neutral in the scale of the Debye length $\lambda_D$, which is an important plasma parameter and will be elucidated in the later section.

In classical plasma theory, three criteria must be satisfied to distinguish plasmas from ionized gases \citep{chen2016}. First, the quasi-neutrality requires that the Debye length $\lambda_D$ is much smaller than the system scale $L$. Second, to make the quasi-neutrality a statistically valid concept, the number $N_D$ of particles in a ``Debye sphere ", namely $N_D := n4/3\pi\lambda_D^3$ with $n$ being the number density of particles, must be large enough. The third criterion states the dominance of the electromagnetic forces over the hydrodynamics forces in driving the motions of particles.

\subsection{Fluid model}
To describe the dynamics of plasmas, a natural and extremely precise way is to treat it as a collection of particles and to monitor the positions and velocities of each particle, namely \emph{N-body dynamics}. It has been proven powerful in analytically describing effects relevant to plasmas \citep{escande2015}.  However, a typical plasma density might be $10^{18}$ particles per $m^3$ \citep{chen2016}, which is way far from affordable for today's computers if one attempts to numerically simulate it. Therefore, relatively crude models are necessary for plasma simulations. The \emph{kinetic model} is another successful tool to describe plasmas. It compiles the state of a collection of particles in the form of probabilistic distributions of velocity and usually consists of the Vlasov-Maxwell equations with velocity distribution functions as variables. See \cite{chaudhary2018} for details on the kinetic description of plasmas.

An even simpler model, \emph{fluid model}, described by the Euler-Maxwell equations, can be derived by taking the moments of the Vlasov-Maxwell equations supplemented with some closure assumption. Specifically, we take the velocity distribution to be $Maxwellian$, by which the distribution can be characterized by its moments. In this way, the fluid model describes the evolution of the moments, e.g. number density, drift velocity, pressure, rather than the distribution. Such a treatment reduces the complexity of the system at the cost of some information. This bridging between the fluid model and the kinetic model applies reasonably for ordinary hydrodynamics where the Maxwellian velocity distribution results from frequent collisions. It seems problematic for a plasma since generally, the collision rate is very low in a plasma due to its dilute nature. However, the experimental work of \cite{langmuir1925} demonstrates the surprising validity of the Maxwellian assumption for plasmas. This phenomenon is known as \emph{Langmuir paradox} which remains unsolved. Readers are referred to \cite[][p. 63]{chen2016} for some discussions. It is worth noting that the fluid model suffices for the majority of plasma phenomena, but certain situations exist where neither the fluid model is applicable, nor is the kinetic theory.

In the fluid model of plasma, each species of particles is equipped with a full set of fluid equations into which the Lorentz force and other plasma effects are incorporated. Maxwell's equations govern how an electromagnetic field is generated by the charged fluids, their motions, and changes of the fields. For such a coupled system, various assumptions are usually made to reduce complexity according to specific scenarios. In the simplest case, one considers a fully-ionized plasma and treats the ions, which are about $10^4$ times heavier than electrons, to be immobilized. Such simplification makes sense when phenomena at the scale of electron motions are of particular interest, and thus leads to so-called \emph{one-fluid} model\footnote{In this report, we think of ions, electrons, and neutral atoms as `fluid', and each `fluid' is described by a set of fluid equations. Also, `species' and `fluid' are synonymous here. Notice that in some literature they have different meanings.} in the sense that only one set of fluid equations is sufficient. To study more realistic situations, motions of multiple species of particles ought to be simulated. Furthermore, other effects like viscosity, ionization, recombination, and collisions could play important roles in many applications.

\subsection{Quasi-neutrality} \label{sec:quasi-neutrality}
Let's imagine such a situation: in a neutral\footnote{It means the net charge vanishes everywhere.} plasma with temperature $T_0$ and electron number density $n_0$, the electrons contained in cube volume $d^3$ are, due to thermal motions, separated from ions. The non-zero charge thus gives rise to an induced electric field, and the energy required for this movement is of magnitude $n_0e^2d^2 / \varepsilon_0$ for a single particle. In the absence of external forces, this energy must come only at the expense of the thermal energy \citep{frank_1972}, which is of magnitude $k_BT_0$. Only when the energy required for the displacement is less than the thermal energy can this local unbalance of charges happen. By balancing these two energy magnitudes we get the length scale

\begin{equation*}
    d \sim \left(\frac{\varepsilon_0k_BT_0}{e^2n_0}\right)^{1/2} =: \lambda_D
\end{equation*}
which is termed \emph{Debye length}\footnote{The Debye length is usually defined for electrons. In fact, one can define different Debye lengths for other species.}. The Debye length is a fundamental property of nearly all plasmas of interest \citep{gibbon2020}. Being a crucial plasma parameter, it is a characteristic length scale over which the perturbation of electric charges is screened due to random thermal motions. With a small $\lambda_D$, the mismatch of positive and negative charges would be limited to a small area. Hence the plasma is approximately \emph{neutral}, namely canceling out in net charges, over a relatively large spatial scale. Such phenomenon of effectively shielding the electrostatic field within a distance of the order $\lambda_D$ is called \emph{Debye shielding}.

Now comes the illustration of \emph{quasi-neutrality}, which is a common property of plasmas. Consider $L_0$ as the characteristic length scale of the system, upon which the dimensionless Debye length $\lambda$ is readily defined as
\begin{equation*}
    \lambda := \frac{\lambda_D}{L_0}.
\end{equation*}
When $\lambda_D$ is much smaller than $L_0$, any local charge unbalance is shielded out within distance of the order $L_0$. The quasi-neutrality refers to such a situation that the plasma is neutral enough such that one can take $\rho \approx 0$, but not so neutral that all the interesting electromagnetic forces vanish \citep{chen2016}. In other words, the quasi-neutrality does not render the plasma a neutral gas such that it has no interaction with the electromagnetic field.

From a mathematical viewpoint, the quasi-neutrality assumption causes the degeneracy of Gauss's law for the electric field to the quasi-neutral condition $\rho = 0$. Despite the vanishing electric charge $\rho$, the electric field is \emph{not} divergence-free. Such degeneracy of the system poses challenges when one has to deal with the non-neutral and quasi-neutral models---two systems of different mathematical properties---at the same time.

\section{Asymptotic-preserving schemes}
 Many important partial differential equations (PDEs) systems involve asymptotic parameters (denoted by $\lambda$ from now on), which might take value from $\lambda = \mathcal{O}(1)$ to $\lambda \ll 1$ at different regimes of the targeting problem, e.g. the relaxation of collision term (or Knudsen number \textbf{Kn}) in Boltzmann equations, Mach number \textbf{Ma} in the Navier-Stokes (N-S) equations, and the Debye length $\lambda_D$ in the Euler-Maxwell system. These parameters have close connections with the characteristic scale of the system, e.g. wave speed and relaxation time. As the asymptotic parameter $\lambda$ approaches zero, the limit solution (if exists) is described by a set of PDEs of a different type from the non-limit one. A \emph{singular perturbation} problem hence arises (See \cite{mark_1995} for related theories). For instance, the Boltzmann equation tends to the compressible N-S equations as \textbf{Kn} $\rightarrow 0$; the compressible N-S equations approach the incompressible case as \textbf{Ma} $\rightarrow 0$, and the fluid pressure switches its hyperbolic character to the elliptic one. To give an intuition about the singular perturbation, consider a toy model
 \begin{equation} \label{equ:toy-model}
     \lambda u' + u = 0, u:[0,\infty]\rightarrow\mathbb{R},
 \end{equation}
 which is a first-order ordinary differential equation (ODE). Let $\lambda$ go to zero, the (pointwise) limit solution satisfies 
 \begin{equation*}
     u = 0,
 \end{equation*}
 which is no longer a first-order ODE.

 Following \cite{jin_2010}, we refer to the original system ($\lambda = \mathcal{O}(1)$) as \emph{microscopic} model while the limiting system ($\lambda \rightarrow 0$) as \emph{macroscopic} model, denoting them by $P^\lambda$ and $P^0$ respectively. If either the macroscopic or the microscopic model is uniformly valid in the whole domain, one can devise schemes accordingly. In fluid dynamics, for example, different solvers are tailored for compressible and incompressible flows. Applying schemes for $\lambda = \mathcal{O}(1)$ to the problem with $\lambda \ll 1$ would be problematic because schemes for $P^\lambda$ are usually subject to CFL-like conditions dependent on $\lambda$. For instance, \textbf{Ma} is inversely proportional to the sound speed; Debye length is inversely proportional to the wave speed. Under-resolving these small scales usually lead to instability. Therefore, $\lambda \ll 1$ would pose very strict restriction on time-step size, leading to prohibitive computational cost. Look back to the toy model (\ref{equ:toy-model}). A reasonable scheme for $\lambda = \mathcal{O}(1)$ is the explicit Euler
 \begin{equation*}
     \lambda\frac{u^{n+1}-u^{n}}{\delta} + u^n = 0.
 \end{equation*}
Yet obviously if the time-step size $\delta$ is fixed, it does not apply to arbitrarily small $\lambda$ due to the stability condition $\delta < 2\lambda$.

 The standard strategy to handle problem $P^\lambda$ with $\lambda \ll 1$ is to solve $P^0$ instead of $P^\lambda$ under the assumption that $P^0$ is a good approximation of $P^\lambda$. For instance, the fluid limit of the Boltzmann equation removes the restrictive stability condition that is dependent on a small relaxation length; the incompressible N-S equations eliminate the need to resolve the sound speed whereas it is necessary in the compressible case. However, challenges occur when sub-domains with $\lambda = \mathcal{O}(1)$ and $\lambda \ll 1$ coexists in the computational domain. To mention some examples, the reentry problem of space shuttles involves both rarefied and dense gases; the plume of a convergence-divergence nozzle experiences \textbf{Kn} from small to very large value \citep{jenny2015}; the plasma sheath near the wall area violates the quasi-neutral assumption in contrast to the quasi-neutral bulk area. To handle such hybrid problems, the classical approach is to split the domain and apply schemes accordingly. However, the criterion for domain splitting is not straightforward and quite empirical. In addition, the projection of solutions from one model to another in the transitional area adds to complexity \citep{jin_2010, degond_2011}. In other words, the non-uniformity of numerical schemes on microscopic and macroscopic models poses a great challenge in handling problems in a transitive regime.

 This dilemma spurs tremendous interests in developing schemes that work \emph{uniformly} well for $\lambda \in [0,1]$ with fixed spatial/time step. Such a scheme is expected to produce satisfactory results without resolving the smallest scale and provide a more elegant and unified solution to problems in a transitive regime. At this point, we introduce the concept of \emph{asymptotic-preserving} (AP) schemes. We resort to the celebrated commuting diagram Fig. \ref{fig:illustration_ap} in \cite{jin_1999} which sketches the property that an AP scheme should possess. As is defined above, $P^\lambda$ and $P^0$ stand for the (solutions of) microscopic and macroscopic models respectively. We introduce the discrete model $P^\lambda_h$ based on some AP scheme, whose solution depends on both the asymptotic parameter $\lambda$ and the spatial (or temporal) step size $h$. We stress that each limit in Fig. \ref{fig:illustration_ap} is attained by letting one parameter go to zero while fixing the other. As a basic requirement, the discrete microscopic model $P^\lambda_h$ should at least converge to its analytic solution $P^\lambda$ as $h\rightarrow0$. The AP property additionally allows the convergence of $P^\lambda_h$ to the limit discrete model $P^0_h$ as $\lambda\rightarrow0$. To accomplish this aim, the stability restriction on $h$ must be independent of $\lambda$. We refer to such property as \emph{asymptotic stability}, following \cite{degond_2011}. Furthermore, the discrete macroscopic model $P^0_h$ must again converge to the macroscopic model $P^0$ as $h\rightarrow0$. Such property is termed \emph{asymptotic consistency}. In other words, a discrete model (scheme) $P^\lambda_h$ is AP if it satisfies the commuting diagram Fig.\ref{fig:illustration_ap}. To make it intuitive, AP schemes preserve the asymptotic limits from the microscopic to the macroscopic models, \emph{at the discrete level} \citep{jin_2010}.

  \begin{figure}
     \centering
     \begin{tikzpicture}
     \node (nw) {\Large $P^\lambda_h$};
     \node (ne) [right=3cm of nw]{\Large $P^\lambda$};
     \node (sw) [below=of nw]{\Large $P^0_h$};
     \node (se) [right=3cm of sw]{\Large $P^0$};
     \draw[->] (nw.east)  -- node [above,midway] {$h\rightarrow0$} (ne.west) ;
     \draw[->] (nw.east)  --  (ne.west) ;
     \draw[->] (nw.south) -- node [left,midway] {$\lambda\rightarrow0$} (sw.north);
     \draw[->] (nw.south) -- node {} (sw.north);
     \draw[->] (sw.east)  -- node [above,midway] {$h\rightarrow0$}(se.west);
     \draw[->] (ne.south) -- node [left,midway] {$\lambda\rightarrow0$} (se.north);
     \end{tikzpicture}
     \caption{Illustration of AP schemes}
     \label{fig:illustration_ap}
 \end{figure}

 We present the definition in a formal way: 
 \begin{definition} \label{def:ap}
 Consider the microscopic model $P^\lambda$ depending on an asymptotic parameter $\lambda$, and assume the macroscopic model $P^0$ exists as the asymptotic limit as $\lambda \rightarrow 0$. The discrete model $P^\lambda_h$, depending both on $\lambda$ and $h$, is \emph{asymptotic-preserving} if
 \begin{itemize}
     \item $P^\lambda_h$ converges to $P^\lambda$ as $h\rightarrow0$ with stability conditions independent of $\lambda$ \emph{(asymptotic stability)};
     \item the asymptotic limit model $P^0_h$ is consistent with $P^0$ \emph{(asymptotic consistency)}.
 \end{itemize}
 \end{definition}
 Looking back again to the model (\ref{equ:toy-model}), one can easily verify that the implicit Euler
 \begin{equation*}
     \lambda\frac{u^{n+1}-u^{n}}{\delta} + u^{n+1} = 0
 \end{equation*}
 is AP.

 Extra properties are also desirable in some cases. In kinetic models with relaxations \citep{jin_2010}, for example, the scheme that can drive \emph{any} initial data into the local equilibrium \emph{in one step} is referred to as \emph{strongly AP}. Instead, if the local equilibrium is reached after long enough, it is referred to as \emph{relaxed AP}. If the initial data is well-prepared and the equilibrium remains for good, it is \emph{weakly AP}. Yet in this thesis, we are only concerned about the essential part of AP property that is specified in Definition \ref{def:ap}.

 Note that the existence of a macroscopic model $P^0$ as the asymptotic limit of microscopic model $P^\lambda$ is presumed. Rigorous proofs of some cases are available, for instance, in \cite{friedman_1968} for general linear PDEs, \cite{klainerman_1981} for the incompressible limit of the Euler equations, \cite{chen_1994, natalini_1996} for conservation laws with stiff relaxation term; \cite{remi_2014} for quasi-neutrality limit in the Euler-Maxwell system. This topic is beyond the scope of this thesis and we just take for granted the existence of such limit models.

 The concept of AP schemes can be traced back to \cite{jin_1995} in which the authors approximate a conservation law by a hyperbolic system with relaxation, and manage to construct a scheme that is stable for an arbitrarily small relaxation parameter. The term ``asymptotic-preserving" was formally introduced in \cite{jin_1999} for handling the diffusive regime of kinetic models. This concept stemmed from and remains very active in kinetic models with stiff relaxations, see \cite{jin_2010} and the references therein. Nonetheless, considerable effort has been put into developing AP schemes for many other problems that admit asymptotic behavior. For instance, in fluid dynamics, the so-called \emph{all-speed} schemes are advantageous when dealing with both compressible and incompressible regimes simultaneously. \cite{degond_2007, haack_2010} devise all-speed schemes for the isentropic Euler equations through fast-slow decomposition of velocity fields. Concerning plasma physics, a number of papers are available dealing with the quasi-neutral limit of the Euler-Poisson (or Euler-Maxwell) system \citep{crispel_2005, degond_2008, degond_2011, degond_2012} and Vlasov-Poisson (or Vlasov-Maxwell) system \citep{degond_2017, degond2017b, zhu_2017}. Besides, \cite{caflisch_1997, berthon2011} concern AP schemes for generic hyperbolic systems with relaxation source terms. For a comprehensive review on AP schemes, readers are referred to \cite{jin_2010} for kinetic and hyperbolic equations, and \cite{degond_2017} for multiscale plasma simulations.

 To devise an AP scheme, one mostly has to consider which terms need to be evaluated implicitly to meet the requirements of the AP property \citep{degond_2011} since stronger stability is usually achieved through implicit treatment in time discretization. We want to stress that unconditionally-stable (with respect to $\lambda$) schemes are not necessarily AP \citep{degond_2011,caflisch_1997}\footnote{It means the asymptotic stability does not necessarily lead to asymptotic consistency as is defined in definition \ref{def:ap}.}. Furthermore, a fully-implicit scheme is very likely to possess AP property, but the computational difficulties arising from solving massive linear or possibly nonlinear equations is potentially a huge cost. As is summarized by \cite{degond_2017}, AP schemes aim to bridge several sets of equations describing the system in different regimes, rather than to alleviate the stability constraints of the numerical methods. In this sense, a fully-implicit scheme is likely to be overkill. We want to achieve the AP property without incurring too much extra cost, which necessitates a smart construction.

 A general methodology for devising AP schemes is suggested by \cite{degond_2012, degond_2017}. They summarize it into two steps. The first step is a ``reformulation" step, where an equivalent form $R^\lambda$ of $P^\lambda$ is to be found such that the $R^\lambda$ explicitly appears as a perturbation of $P^0$. Afterward, one discretizes $P^\lambda$ by a scheme $P^\lambda_h$ with a certain amount of implicitness in such a way that the reformulation procedure can be performed at the discrete level, and leads to a reformulated discrete model $R^\lambda_h$ that is consistent to $R^\lambda$. However, this procedure is somehow heuristic, and its applicability to other perturbation problems remains questionable. It is also infeasible to rigorously prove the AP property for a given scheme, which is largely owing to the fuzzy notations of ``convergence" of numerical schemes. Even the linearized stability, as a necessary condition, is difficult to analyze in a multi-dimensional framework. Therefore, for the time being, we can state the AP property and devise the AP scheme merely in a heuristic way, and mostly has to resort to numerical experiments for verification.

 To summarize, the advantage of AP schemes over classical ones is twofold: for problems involving different regimes, the macroscopic regime ($\lambda \ll 1$) does not suffer from the strict stability requirement whereas the microscopic regime ($\lambda \sim \mathcal{O}(1)$) can be resolved as usual; the scale of interest, namely the resolution that we desire, is free to choose regardless of the value of the asymptotic parameter that represents the scale of the system.

\chapter{Discussion on Asymptotic-preserving Property}

On portraying the AP property through the commuting diagram Fig. \ref{fig:illustration_ap} and Definition \ref{def:ap}, we convey the principle of how an AP scheme is supposed to behave. Yet these illustrations are somehow given in general terms and lack precise statements. Another concern lies in the potential entanglement between the two requirements of AP schemes, namely asymptotic stability and asymptotic consistency Since these two statements are very strong, decomposing them into weaker ones is likely to help identify essential components to yield AP property. Due to the shortage of theoretical analysis on AP schemes, we humbly make some attempts here.

In this chapter, the AP property, as well as the related arguments, are rephrased in a relatively rigorous way. Specifically, we want to explore the connection of the four asymptotic relations depicted in the commuting diagram Fig. \ref{fig:illustration_ap}, relying on which we try to present an argument about a set of sufficient conditions that lead to an AP scheme.

Based on the formulation the commuting diagram Fig. \ref{fig:illustration_ap}, we rephrase the AP property by a uniform convergence statement: a discrete model $P^\lambda_h$ is AP if and only if
\begin{equation} \label{arg:ap}
    \exists C, k > 0,\ \ \  \|P^\lambda_h - P^\lambda\| \leq Ch^k,\ \ \ \forall \lambda \in [0,1]  
\end{equation}
where $\|\cdot\|$ represents some suitable norm.
We repeat the definitions: the asymptotic parameter $\lambda$; the discretization parameter $h$, e.g., time step size and cell size; the (solution of) discrete model $P^\lambda_h$; the (solution of)  continuous model $P^\lambda$. This amounts to stating a scheme that converges to the analytical solution uniformly well for arbitrary $\lambda$ and especially for the macroscopic model $P^0$, which is nothing but the philosophy of AP schemes.

Next, we present an argument that relates the convergence relationships in the commuting diagram with the above uniform convergence statement.
\begin{proposition} \label{prop:1}
Assume $P^0_h$ is well-defined. If the following conditions are satisfied
\begin{itemize}
    \item[$(*)$] $P^\lambda_h \xrightarrow[]{h\rightarrow0}P^\lambda$ in the sense that $\exists \alpha, \Tilde{\alpha}, C_1$ being constant independent of $h, \lambda$ such that
    \begin{equation*}
        \|P^\lambda_h - P^\lambda\| \leq C_1\frac{h^\alpha}{\lambda^{\Tilde{\alpha}}},\ \ \ \forall \lambda \in (0,1],
    \end{equation*}
    \item[$(**)$] $P^\lambda_h \xrightarrow[]{\lambda\rightarrow0} P^0_h$ in the sense that $\exists \beta, C_2$ being constant independent of $h, \lambda$ such that
    \begin{equation*}
        \|P^\lambda_h - P^0_h\| \leq C_2\lambda^\beta,\ \ \ \forall h > 0,
    \end{equation*}
    \item[$(\$)$] $P^\lambda \xrightarrow[]{\lambda\rightarrow0}P^0$ in the sense that $\exists C_3, \gamma$ being constant independent of $h, \lambda$ such that
    \begin{equation*}
        \| P^\lambda - P^0\| \leq C_3\lambda^\gamma, \ \ \ \forall \lambda \in [0,1],
    \end{equation*}
\end{itemize}
Then we have $P^\lambda_h \xrightarrow[]{h\rightarrow0} P^\lambda$ uniformly on $\lambda$, namely argument (\ref{arg:ap}).
\end{proposition}
\begin{proof}
Let $\epsilon$ be a constant satisfying $\alpha - \epsilon > 0$. For any $\lambda \in [0,1]$, we have the following two situations.

If $\lambda^{\Tilde{\alpha}} < h^{\alpha - \epsilon}$, we pick $\Tilde{\lambda} = h^\frac{\alpha - \epsilon}{\Tilde{\alpha}}$, and we have
\begin{align*}
    \|P^\lambda_h - P^\lambda\| &\leq \|P^\lambda_h - P^{\Tilde{\lambda}}_h\| + \|P^{\Tilde{\lambda}}_h - P^{\Tilde{\lambda}}\| + \|P^{\Tilde{\lambda}} - P^\lambda\| \\
    & \leq C_2(\lambda^\beta + \Tilde{\lambda}^\beta) + C_1\frac{h^\alpha}{{\Tilde{\lambda}}^{\Tilde{\alpha}}} + C_3(\Tilde{\lambda}^\gamma + \lambda^\gamma) \\
    & \leq 2C_2h^{\frac{\alpha - \epsilon}{\Tilde{\alpha}}\beta} + C_1h^\epsilon + 2C_3 h^{\frac{\alpha - \epsilon}{\Tilde{\alpha}}\gamma} \\
    & \leq 5\text{max}\{C_1,C_2,C_3\}h^{\text{min}\{\frac{\alpha - \epsilon}{\Tilde{\alpha}}\beta, \epsilon, \frac{\alpha - \epsilon}{\Tilde{\alpha}}\gamma\}}.
\end{align*}
The triangle inequality is applied to the first and second inequalities.

If $\lambda^{\Tilde{\alpha}} \geq h^{\alpha - \epsilon}$, we can directly apply condition $(*)$, which results in
\begin{align*}
    \|P^\lambda_h - P^\lambda\| \leq C_1\frac{h^\alpha}{\lambda^{\Tilde{\alpha}}} \leq C_1h^\epsilon.
\end{align*}
Combine these two situations, and we can conclude.
\end{proof}
Condition $(*)$ amounts to saying that the scheme converges to the right solution only when the discretization parameter $h$ resolves $\lambda$ in some way. Otherwise, the scheme diverges. This should be easily met by any classical scheme which is not necessarily AP. The asymptotic stability is enforced in Condition $(**)$ which implies that for a fixed $h$, no $\lambda$ should incur instability. Being the underlying assumption, the convergence of the microscopic model to the macroscopic model, as is depicted by Condition $(\$)$, is always taken for granted throughout the thesis. Therefore, Condition $(*)$ and $(\$)$ are usually satisfied by a suitable problem and a classical scheme. Proposition \ref{prop:1} hence provides some guidance on how to design an AP scheme, that is, to incorporate Condition $(**)$ into the scheme.

In the sequel, we consider a single-step scheme in the form $u_{\lambda,h}^{m+1} = \Phi_{\lambda,h}^m u_{\lambda,h}^{m}$, where $\Phi_{\lambda,h}^m$ is the discrete evolving operator, and $u_{\lambda,h}^{m}$ is the numerical solution at $m$-th time step. Instead of directly arguing the convergence of solutions, as Proposition \ref{prop:1}, starting from such a scheme allows us to have finer control of their properties, especially the stability and consistency. The conclusion based on the scheme is more revealing in understanding AP properties and more instructive in the construction of AP schemes.

\begin{proposition}{\emph{(A sufficient argument for AP schemes)}} \label{prop:2}

Consider a suitable norm $\|\cdot\|$ and a normed space $\mathcal{V}$ as the solution space. For a singular perturbation problem $P^\lambda$, assume Condition $(\$)$ in Proposition \ref{prop:1} is satisfied, and the solutions are uniformly bounded, that is, $\|\Tilde{u}_\lambda\| \leq C_u,\ \forall \lambda \in [0,1].$
For a single-step scheme $u_{\lambda,h}^{m+1} = \Phi_{\lambda,h}^m u_{\lambda,h}^{m}$, assume $\Phi_{0,h}^m$ is well-defined. We denote the exact solution by $\Tilde{u}_{\lambda,h}^m$. If the scheme satisfies the following conditions
\begin{itemize}
    \item[\emph{(AS)}]  $\Phi_{\lambda,h}(\cdot)$ is $(1+hC_\Phi)$-Lipschitz $\forall\lambda\in[0,1]$  where the constant $C_\Phi$ does not depend on $\lambda, h, m$;

    \item[\emph{(C)}] the non-limit scheme is \emph{consistent} in the sense that 
    \begin{equation*}
        \| \Tilde{u}^{m+1}_{\lambda,h} - \Phi_{\lambda,h}^m\Tilde{u}^{m}_{\lambda,h}\| \leq C_c\frac{h^2}{\lambda}, \ \ \ \forall \lambda \in (0,1],
    \end{equation*}
    where the constant $C_c$ does not depend on $\lambda, h, m$;

    \item[\emph{(AC)}] the limiting scheme is \emph{consistent} in the sense that 
    \begin{equation*}
        \| \Tilde{u}^{m+1}_{0,h} - \Phi_{0,h}^m\Tilde{u}^{m}_{0,h}\| \leq \Tilde{C_c}h^2, 
    \end{equation*}
    where the constant $\Tilde{C_c}$ does not depend on $h, m$;

    \item[\emph{(L)}] $\Phi_{\lambda,h}^m$ is Lipschitz with respect to $\lambda$ in the sense that
    \begin{equation*}
        \| (\Phi_{\lambda,h}^m - \Phi_{0,h}^m)u \| \leq \Tilde{C}_\Phi\|u\|h\lambda,\ \ \ \forall u \in \mathcal{V}, 
    \end{equation*}
    where the constant $\Tilde{C}_\Phi$ does not depend on $h$,
\end{itemize}
then the scheme is AP in the sense of (\ref{arg:ap}).
\end{proposition}
\begin{proof}[Proof]
Providing Proposition \ref{prop:1}, it suffices to check Condition $(*)$ and $(\$)$.

>>>>

\documentclass{report}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsthm}

\title{APPM report - Draft}
\author{T.W. Yu, R. Fuchs, R. Hiptmair}

\begin{document}
\maketitle

\chapter{Introduction}

\end{document}
